### 2025-07-31 mythbusters.qmd
20250720 - conclusion sections
20250731 - research question, hypothesis, and prediction

Let's examine the question empirically.
In order to do that, we have to make some assumptions/constraints.
In this example, we will make the following constraints:

@@

There is a widely held belief that NFL players perform better in the last year of the contract because they are motivated to gain another contract.
There is some evidence in the NBA and MLB that players tend to perform better in their contract year.
We evaluated this possibility among NFL players who were Quarterbacks, Running Backs, Wide Receivers, or Tight Ends.
We evaluated a wide range of performance indexes, including Quarterback Rating, yards per carry, points added, expected points added, receiving yards, and fantasy points.
None of the positions showed significantly better performance in their contract year for any of the performance indexes.
By contrast, if anything, players tended to perform more poorly during their contract year, as operationalized by fantasy points, receiving yards (WR/TE), and EPA from receiving plays (WR/TE), even when controlling for player and age experience.
In sum, we did not find evidence in support of the contract year hypothesis and consider this myth debunked.
However, we are open to this possibility being reexamined in new ways or with additional performance

>>>>

Let's examine the question empirically.
Our <u>research questions</u> is: Do players perform better in their "contract year" (i.e., the last year of their contract)?
Our <u>hypothesis</u> is that players are motivated to get larger contracts (more money), leading players in their contract year to try harder and perform better.
If the hypothesis is true, we <u>predict</u> that players who are in their contract year will tend to score more fantasy points than players who are not in their contract year.

In order to test this question empirically, we have to make some assumptions/constraints.
In this example, we will make the following constraints:

@@

There is a widely held belief that NFL players perform better in the last year of the contract because they are motivated to gain another contract.
There is some evidence in the NBA and MLB that players tend to perform better in their contract year.
We evaluated this possibility among NFL players who were Quarterbacks, Running Backs, Wide Receivers, or Tight Ends.
We evaluated a wide range of performance indexes, including Quarterback Rating, yards per carry, points added, expected points added, receiving yards, and fantasy points.
None of the positions showed significantly better performance in their contract year for any of the performance indexes.
By contrast, if anything, players tended to perform more poorly during their contract year, as operationalized by fantasy points, receiving yards (WR/TE), and EPA from receiving plays (WR/TE), even when controlling for player and age experience.
In sum, we did not find evidence in support of the contract year hypothesis and consider this myth debunked.
However, we are open to this possibility being reexamined in new ways or with additional performance metrics.
